---
title: "Pain 01 Data Structure"
author: "Jack VanSchaik"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, results="hide", message=FALSE}
library(fst)
library(ggnetwork)
library(jsonlite)
library(network)
library(quanteda)
library(tidyverse)
```

## Read In Data

```{r}
gene_names <- read_fst("scratch/gene_names.fst")
pm_df_2 <- read_fst("scratch/pm_df_2.fst")
```

## Planning

I think we'll probably want three json objects:

* Gene Names: List of gene names we want to query
* 

Something we'll need for sure:

* pre-computed term frequency for each gene
* for proper IDF, we will need to tokenize

## Text Processing

#### Get Vector of Gene Names

```{r}
genes <- gene_names$gene_name
```

#### Create Corpus and Tokenize

```{r}
pm_df_2 %>%
    corpus(docid_field="pubmed_id", text_field="abstract") %>%
    tokens(
        what="word",
        remove_punct=TRUE,
        remove_symbols=TRUE
    ) %>%
    tokens_tolower ->
    pm_tok
```

#### Get Gene Counts

```{r}
pm_tok %>%
    dfm %>%
    dfm_keep(pattern = genes, valuetype = "fixed") %>%
    convert(to = "data.frame") ->
    pm_tf
```

```{r}
pm_tf %>%
    pivot_longer(-doc_id) %>%
    filter(value != 0) %>%
    select(
        gene = name,
        pubmed_id = doc_id,
        tf = value
    ) %>%
    arrange(gene, desc(tf)) ->
    pm_gene_docs
```

## Prep Data For Network Plot

```{r}
pm_gene_docs %>%
    inner_join(pm_gene_docs, by="pubmed_id") %>%
    arrange(pubmed_id) %>%
    filter(gene.x != gene.y) %>%
    select(gene.x, gene.y) %>%
    distinct %>%
    network ->
    pm_net

pm_net %>%
    ggnetwork(cell.jitter=0.75) %>%
    ggplot(aes(x = x, y = y, xend = xend, yend = yend, label=vertex.names)) +
    geom_edges(color = "grey50") +
    geom_nodetext() + 
    theme_blank()

```


## Create json objects

#### TF json object

```{r}
gene_doc_sp <- split(pm_gene_docs, as.factor(pm_gene_docs$gene))
gene_doc_L <- lapply(gene_doc_sp, function(j) as.list(select(j, pubmed_id, tf)))
names(gene_doc_L) <- names(gene_doc_sp)

gene_doc_json <- jsonlite::toJSON(gene_doc_L)
```
## Export Data

```{r}
write_file(gene_doc_json, "web_data/gene_doc.json")
write_rds(pm_net, "scratch/pm_net.rds")
```

